---
title: "Penguin Size Comparisons"
author: "Anastasia Kunz"
date: "11/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effsize)
library(broom)
library(ggbeeswarm)
library(car)
```


# Part 1. T-tests

Mass comparison between male and female chinstrap penguins

```{r}
chinstrap <- penguins %>% 
  filter(species == "Chinstrap")

# Histograms

ggplot(chinstrap, aes(x = body_mass_g))+
  geom_histogram(bins = 10)+
  facet_wrap(~sex)
#  somewhat normal/ bell shaped mean should be okay. male means seem a bit larger than female means.

# Quantile- Quantile

ggplot(chinstrap, aes(sample = body_mass_g))+
  geom_qq() +
  facet_wrap(~sex)

# Descriptive Stats

chinstrap %>% 
  group_by(sex) %>% 
  summarize(
    mean_mass = mean(body_mass_g),
    median_mass = median(body_mass_g),
    sd_mass = sd(body_mass_g),
    sample_size = n()
  )

```


# T test

Get vectors of body mass for female and male chinstraps separatley.

```{r}
chinstrap_f <- chinstrap %>% 
  filter(sex == "female") %>% 
  pull(body_mass_g)

chinstrap_m <- chinstrap %>% 
  filter(sex == "male") %>% 
  pull(body_mass_g)
```

Now, can run t-test:

1. Is there a significant difference in body mass between male and female chinstrap penguins?
--> not are males bigger!

```{r}
chinstrap_ttest <- t.test(chinstrap_f, chinstrap_m)
chinstrap_ttest

# effect size in effsize

cohen.d(chinstrap_f, chinstrap_m)
```

Are male chinstrap penguins significantly larger in body mass than female chinstraps?
--> one sided

```{r}
t.test(chinstrap_m, chinstrap_f, alternative = "greater")
# must be lined up; this says are males greater than female? instead of the other way around
```

# Part 2. One- way ANOVA and Beeswarm graph

## Data Exploration

1. table of counts by penguin species and sex

```{r}
penguin_count <- penguins %>% 
  count(species, sex)
#have to call it to get it to show up
penguin_count

```


create a subset of only female penguins

```{r}
penguin_f <- penguins %>% 
  filter(sex == "female")

#Hist

ggplot(penguin_f, aes(x = flipper_length_mm))+
  geom_histogram(aes(fill = species))+
  facet_wrap(~species, scales = "free") #each facet can have different scales

#QQ

ggplot(penguin_f, aes(sample = flipper_length_mm))+
  geom_qq(aes(color = species))+
  facet_wrap(~species, scales = "free")

#could fill by species or facet wrap or both

#Summary stats

penguin_stats <- penguin_f %>% 
  group_by(species) %>% 
  summarize(
    mean_flipper_length = mean(flipper_length_mm),
    sd_flipper_length = sd(flipper_length_mm),
    sample_size = n(),
    se_flipper_length = sd(flipper_length_mm)/sqrt(n()),
    var_flipper_length = var(flipper_length_mm)
  )

leveneTest(flipper_length_mm ~ species, data = penguin_f)#var of flip length as a funct of sp
```

## Bee swarm plot and a mean + or - a std dev

```{r}
ggplot()+
  geom_beeswarm(data = penguin_f, 
                aes(x = species, y = flipper_length_mm),
                color = "grey70",
                size = 3,
                alpha = 0.5) +
  geom_point(data = penguin_stats,
             aes(x = species, y = mean_flipper_length),
             size = 4)+
  geom_errorbar(data = penguin_stats,
                aes(x = species,
                  ymax = mean_flipper_length + sd_flipper_length,
                    ymin = mean_flipper_length - sd_flipper_length),
                width = 0.1) + 
  theme_light()
```

## One Way Anova

```{r}
penguin_aov <- aov(flipper_length_mm ~ species, data = penguin_f)
summary(penguin_aov)


#Post hoc test

TukeyHSD(penguin_aov)

```

for unbalanced design
type 3 anova with car package

```{r}
penguin_lm <- lm(flipper_length_mm ~species, data = penguin_f)
penguin_aovt3 <- Anova(penguin_lm, type = "III")
penguin_aovt3
```

unless crazy unbalanced then they will be pretty close.

